
\documentclass{article}
\usepackage{import}
\usepackage{epstopdf}
\usepackage{bsymb}
\usepackage{alltt}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{multicol}
\usepackage{color}
\usepackage{hyperref}

\begin{document}

\title{Tasking Event-B: A Code Generation Extension for Event-B}


\author{Andrew Edmunds\\
School of Electronics and Computer Science,\\
University of Southampton, UK\\
\url{ae2@ecs.soton.ac.uk}\\
\and
Michael Butler\\
School of Electronics and Computer Science,\\
University of Southampton, UK\\
\url{mjb@ecs.soton.ac.uk}\\
}

\maketitle
%------------------------------------------------------------------------------
% Abstract
%
\begin{abstract}
Event-B is a formal approach to modelling systems, based on set theory, predicate logic and arithmetic. Various techniques are used during modelling, including refinement, and decomposition. This article describes an extension to the Event-B approach, which we call Tasking Event-B, that facilitates automatic generation of source code, from annotated Event-B models. We believe that automatic code generation makes a useful contribution to the Rodin tool-set, by contributing a link in a coherent tool-chain. Automatically generating code from the Event-B model can be seen as a productivity enhancement. It removes a source of errors, that of manually coding for each development. To validate the approach we have undertaken case studies and taken part in an industrial collaboration. We present a number of case-studies to illustrate our work, in this article.
\end{abstract}

\include{Intro}

\include{Event-B}

\include{taskingEB}

\include{tooling}

\include{conc}

%------------------------------------------------------------------------------
% Refs:
%
\label{sect:bib}
\bibliographystyle{plain}
\bibliography{MyBibTex}
\end{document}

% EOF
